\mythesistitle

### 2011-12-01 src/synopsis.tex
spell check done with aspell. however, previous commit was printed and circulated. some typos in there - deal with it.

Section~\ref{sec:background} gives a brief survey of COP and
optimization problems related to it followed by motivation for the
thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
presents a summary of our results on the extention of COP namely, the
tree path labeling problem. \tnote[TH]{update for thesis}

@@

Once a matrix has been detected to not have COP (using any of the COT
algorithms mentioned earlier), it is naturally of interest to find out
the smallest forbidden substructure (in terms of number of rows and/or
columns and/or number of entries that are {\un}s). \cite{d08phd}
discusses a couple of algorithms which are efficient if the number of
{\un}s in a row is small. This is of significance in the case of sparce
matrices where this number is much lesser than the number of
columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
on number of {\un}s in any column but has at most $\Delta$ {\un}s in
any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
similar problems which deals with inducing COP on a matrix. In {\sc
  Min COS-R (Min COS-C)} the question is to find the minimum number of
rows (columns) that must be deleted to result in a matrix with COP.
In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
maximum number of rows (columns) that induces a submatrix with
COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
a submatrix $M'$ with all columns\tnote[GTC]{check if b75 deals with
  COP col or COP row. also is it any submatrix with k less than r rows
  or submatrix must have all columns?} but a maximum cardinality
subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
corrects an error of the abridged proof of this reduction as given in
\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
giving an extensive survey of the previously existing results which
are almost exhaustively all approximation results and hardness
results. Taking this further, \cite{d08phd} presents new results in
the area of parameterized algorithms for this
problem\tnote[TH]{elaborate - what are the results?}.

@@

\cite{jkckv04} discusses the use of matrices with almost-COP (instead
of one block of consecutive {\un}s, they have $x$ blocks, or {\em
  runs}, of consecutive {\un}s and $x$ is not too large) in the
storage of very large databases.  The problem is that of reordering of
a binary matrix such that the resulting matrix has at most $k$ runs of
{\un}s. This is proven to be NP hard using a reduction from the
hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
\tnote[TH]{(1) A connection of COP problem to the travelling salesman
  problem is also introduced. what does this mean? -- COP can be used
  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
version of the $k$-run problem, i.e. minimization of number of blocks
of ones is proven to be NP complete by cite:k77}\tnote[GTC]{are these two
  the same?} \tnote[TH]{what is the reduction?} \tnote[TH]{other
  problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) --
  cite:th98 (generalization of COP - minimax, biotonic column) Tucker}

@@

\def \probdefwidth {0.85\linewidth}
\def \FTPL {{\sc Feasbile Tree Path Labeling}}
\def \CFTPL {\sc Compute Feasible Path Labeling}

@@

%\vspace{3mm}
{\small
\begin{minipage}[h]{5in}
 % \begin{singlespace}
 \vspace{2mm}
  {\large \CFTPLKTREE}\\
  \begin{tabular}[t]{l|l}
    \hline\\
    {\tt Input} & 
    \begin{minipage}[t]{\probdefwidth}
      A hypergraph $\cF$ with vertex set $U$ such that every hyperedge
      $S \in \cF$ is of cardinality atmost $k+2$ and a {\kstar} $T$.\\
    \end{minipage}\\
    {\tt Question} &
    \begin{minipage}[t]{\probdefwidth}
      Does there exist a set of paths $\cP$ from $T$ and a bijection
      $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
        true} on $(\cF, T, \cl)$.
    \end{minipage}\\
  \end{tabular}
  % \end{singlespace}
\end{minipage}\\
}

@@

We also have an algorithm for solving {\CFTPL} with no restrictions
which runs in exponential time. 
% and explores a partial order theory of
% matrices related to {FTPL}.
% gives interesting insight
% into the problem of 
This algorithm finds a path labeling from $T$ by decomposing the
problem into smaller subproblems of finding path labeling of subsets
of $\cF$ from subtrees of $T$. Given the fact that binary matrices
naturally represent a set
system % , we look at the matrix representation of subsets of $\cF$.
and that the overlap relation between the sets involved is an obvious
equivalence relation, $\cF$ naturally partitions into equivalence
classes known as overlap components. In the context of COP, {\em overlap
components} were used in \cite{wlh02} and \cite{kklv10}. Moreoever,
\cite{nsnrs09} discovered that these equivalence classes form a total
order. We extend this to TPL and find that when $\cF$ is a path
hypergraph\footnote{If there exists an FTPL for a hypergraph $\cF$, it
  is called a path hypergraph.}, the classes can be partially ordered
as an in-tree in polynomial time. Once $\cF$ is
``broken'' into overlap components, one must identify the subtree of
$T$ that it needs to map to and this is the hard part which is
currently open to be solved in polynomial time.

@@

We give a characterization for feasible tree path labeling of path
hypergraphs. The proof for this is constructive and computes a
feasiblility bijection mapping vertices of the hypergraph to vertices
of the given target tree. This thesis also discovered an exponential algorithm
that computes a feasible TPL to a given hypergraph if it is a path hypergraph.

@@

Whether TPL on general trees is solvable in $P$ remains open. So do
optimization opportunities in TPL (possible extentions to optimization for COP in the
Section~\ref{sec:optcop}.\tnote[TH]{link to section}).

>>>>

Section~\ref{sec:background} gives a brief survey of COP and
optimization problems related to it followed by motivation for the
thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
presents a summary of our results on the extension of COP namely, the
tree path labeling problem. \tnote[TH]{update for thesis}

@@

Once a matrix has been detected to not have COP (using any of the COT
algorithms mentioned earlier), it is naturally of interest to find out
the smallest forbidden substructure (in terms of number of rows and/or
columns and/or number of entries that are {\un}s). \cite{d08phd}
discusses a couple of algorithms which are efficient if the number of
{\un}s in a row is small. This is of significance in the case of sparse
matrices where this number is much lesser than the number of
columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
on number of {\un}s in any column but has at most $\Delta$ {\un}s in
any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
similar problems which deals with inducing COP on a matrix. In {\sc
  Min COS-R (Min COS-C)} the question is to find the minimum number of
rows (columns) that must be deleted to result in a matrix with COP.
In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
maximum number of rows (columns) that induces a submatrix with
COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
a submatrix $M'$ with all columns\tnote[GTC]{check if b75 deals with
  COP col or COP row. also is it any submatrix with k less than r rows
  or submatrix must have all columns?} but a maximum cardinality
subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
corrects an error of the abridged proof of this reduction as given in
\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
giving an extensive survey of the previously existing results which
are almost exhaustively all approximation results and hardness
results. Taking this further, \cite{d08phd} presents new results in
the area of parameterized algorithms for this
problem\tnote[TH]{elaborate - what are the results?}.

@@

\cite{jkckv04} discusses the use of matrices with almost-COP (instead
of one block of consecutive {\un}s, they have $x$ blocks, or {\em
  runs}, of consecutive {\un}s and $x$ is not too large) in the
storage of very large databases.  The problem is that of reordering of
a binary matrix such that the resulting matrix has at most $k$ runs of
{\un}s. This is proved to be NP hard using a reduction from the
Hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
\tnote[TH]{(1) A connection of COP problem to the travelling salesman
  problem is also introduced. what does this mean? -- COP can be used
  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
version of the $k$-run problem, i.e. minimization of number of blocks
of ones is proven to be NP complete by cite:k77}\tnote[GTC]{are these two
  the same?} \tnote[TH]{what is the reduction?} \tnote[TH]{other
  problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) --
  cite:th98 (generalization of COP - minimax, biotonic column) Tucker}

@@

\def \probdefwidth {0.85\linewidth}
\def \FTPL {{\sc Feasible Tree Path Labeling}}
\def \CFTPL {\sc Compute Feasible Path Labeling}

@@

%\vspace{3mm}
{\small
\begin{minipage}[h]{5in}
 % \begin{singlespace}
 \vspace{2mm}
  {\large \CFTPLKTREE}\\
  \begin{tabular}[t]{l|l}
    \hline\\
    {\tt Input} & 
    \begin{minipage}[t]{\probdefwidth}
      A hypergraph $\cF$ with vertex set $U$ such that every hyperedge
      $S \in \cF$ is of cardinality at most $k+2$ and a {\kstar} $T$.\\
    \end{minipage}\\
    {\tt Question} &
    \begin{minipage}[t]{\probdefwidth}
      Does there exist a set of paths $\cP$ from $T$ and a bijection
      $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns {\bf
        true} on $(\cF, T, \cl)$.
    \end{minipage}\\
  \end{tabular}
  % \end{singlespace}
\end{minipage}\\
}

@@

We also have an algorithm for solving {\CFTPL} with no restrictions
which runs in exponential time. 
% and explores a partial order theory of
% matrices related to {FTPL}.
% gives interesting insight
% into the problem of 
This algorithm finds a path labeling from $T$ by decomposing the
problem into smaller subproblems of finding path labeling of subsets
of $\cF$ from subtrees of $T$. Given the fact that binary matrices
naturally represent a set
system % , we look at the matrix representation of subsets of $\cF$.
and that the overlap relation between the sets involved is an obvious
equivalence relation, $\cF$ naturally partitions into equivalence
classes known as overlap components. In the context of COP, {\em overlap
components} were used in \cite{wlh02} and \cite{kklv10}. Moreover,
\cite{nsnrs09} discovered that these equivalence classes form a total
order. We extend this to TPL and find that when $\cF$ is a path
hypergraph\footnote{If there exists an FTPL for a hypergraph $\cF$, it
  is called a path hypergraph.}, the classes can be partially ordered
as an in-tree in polynomial time. Once $\cF$ is
``broken'' into overlap components, one must identify the subtree of
$T$ that it needs to map to and this is the hard part which is
currently open to be solved in polynomial time.

@@

We give a characterization for feasible tree path labeling of path
hypergraphs. The proof for this is constructive and computes a
feasibility bijection mapping vertices of the hypergraph to vertices
of the given target tree. This thesis also discovered an exponential algorithm
that computes a feasible TPL to a given hypergraph if it is a path hypergraph.

@@

Whether TPL on general trees is solvable in $P$ remains open. So do
optimization opportunities in TPL (possible extensions to optimization for COP in the
Section~\ref{sec:optcop}.\tnote[TH]{link to section}).
